Research Cases

### 2025-10-11 Code/R/Research/cases_research.Rmd
C++ implementation of 1d series reconstruction from tensor

```{r}
library(Rssa)
library(svd)
library(scales)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = getwd())
source(file = "./Source/TSSA.R")
```

@@

```{r}
# no rates
rates.true <- c(0, 0)
plot.name <- "_no_rates"
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# plot.name <- "_small_eq_rates"
# both rates -0.02
# rates.true <- c(-0.02, -0.02)
# plot.name <- "_large_eq_rates"
# rates from paper
# rates.true <- c(-0.01, -0.02)
# plot.name <- ""

freqs.true <- c(0.2, 0.22)
N <- 25

signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(N - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(N - 1))

groups_ssa <- list(1:4)
max.r <- max(sapply(groups_ssa, max))
Ls_ssa <- seq(from = max.r + 1,
              to = (N + 1) %/% 2,
              by = 1)

groups_hossa <- list(1:4)
trunc_dims <- list(1, 1:2, 1:3)
max.r <- max(sapply(groups_hossa, max))

# I >= L >= J
Is <- seq(from = max.r + 1,
          to = N - max.r - 1,
          by = 1)
Ls <- seq(from = max.r + 1,
          to = N - max.r - 1,
          by = 1)

correct.dims <- which(outer(Is, Ls, (\(i, l) ifelse((i >= l) &
                                                      (N - i - l + 2 > max.r) &
                                                      (N - i - l + 2 <= l), TRUE, FALSE
))), arr.ind = TRUE)

@@

tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is[correct.dims[, 1]]
tens.dims[, "L"] <- Ls[correct.dims[, 2]]
tens.dims[, "J"] <- N - tens.dims[, "I"] - tens.dims[, "L"] + 2

@@

sds <- c(0.2, 0.6)

@@

```{r}
R <- 40

ssa_err <- calc_errors_by_params(params_grid_ssa, R, pred_ssa_rec, rmse_ts_nd, signal, seed = 5)
hossa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_rec, rmse_ts_nd, signal, seed = 5)
```

>>>>

```{r}
setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(Rssa)
library(svd)
library(scales)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(Rcpp)
library(RcppArmadillo)
library(future.apply)
library(progressr)

source(file = "./Source/TSSA.R")
```

@@

```{r}
construct_periodics <- function(N,
                                ampl,
                                rate,
                                freq,
                                phase,
                                complex = TRUE,
                                from = 0,
                                by = 1) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")

  if (complex) {
    period_f <- exp
    unit <- 1i
  }
  else {
    period_f <- cos
    unit <- 1
  }

  colSums(ampl * exp(rate) * period_f(phase + 2 * pi * unit * freq %o% seq(
    from = from,
    length.out = N,
    by = by
  )))
}

calc_rank <- function(ampl, rate, freq, phase, complex = TRUE, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")

  (2 - complex) *
    (sum(A != 0) + nrow(unique(matrix(c(freqs.true, rate), ncol = 2))) - length(A)) - sum(abs(freqs.true - 0.5) < tol)
}
```

# Two periodics case
## Short series
```{r}
# no rates
rates.true <- c(0, 0)
rates_case <- "_nr"
# small equal rated
# rates.true <- c(-0.01, -0.01)
# plot.name <- "_ser"
# large equal rates
# rates.true <- c(-0.02, -0.02)
# plot.name <- "_ler"
# different rated
# rates.true <- c(-0.01, -0.02)
# plot.name <- "_dr"

# no freqs
freqs.true <- c(0, 0)
freqs_case <- "_nf"
# different freqs
# freqs.true <- c(0.2, 0.22)
# freqs_case <- "_df"
# close different freqs
# freqs.true <- c(0.02, 0.0205)
# freqs_case <- "_cdf"

# equal amplitudes
A <- c(1, 1)
A_case <- "_ea"
# different amplitudes
# A <- c(1, 2)
# A_case <- "_da"

# equal phases
phases.true <- c(0, 0)
phases_case <- "_ep"
# different phases
# phases.true <- c(0, pi / 3)
# phases_case <- "_dp"

complex.signal <- TRUE
complex_case <- "_c" # complex
# complex.signal <- FALSE
# complex_case <- "_r" # real

full_case <- paste0(rates_case, freqs_case, A_case, phases_case, complex_case)

max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal)

N <- 25
stopifnot(N > max.r * 2 + 1)
signal <- construct_periodics(N, A, rates.true, freqs.true, phases.true, complex.signal)

groups_ssa <- list(1:max.r)
Ls_ssa <- seq(from = max.r + 1,
              to = (N + 1) %/% 2,
              by = 1)

groups_hossa <- list(1:max.r)
trunc_dims <- list(1, 1:3)

# I >= L >= J
Is_tmp <- seq(from = max.r + 1,
          to = N - max.r - 1,
          by = 1)
Ls_tmp <- seq(from = max.r + 1,
          to = N - max.r - 1,
          by = 1)

correct.dims <- which(outer(Is_tmp, Ls_tmp, (\(i, l) ifelse((i >= l) &
                                                      (N - i - l + 2 > max.r) &
                                                      (N - i - l + 2 <= l), TRUE, FALSE
))), arr.ind = TRUE)

@@

tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is_tmp[correct.dims[, 1]]
tens.dims[, "L"] <- Ls_tmp[correct.dims[, 2]]
tens.dims[, "J"] <- N - tens.dims[, "I"] - tens.dims[, "L"] + 2

@@

sds <- seq(0, 0.6, by = 0.2)

@@

```{r}
R <- 500

ssa_err <- calc_errors_by_params(params_grid_ssa, R, pred_ssa_rec, rmse_ts_nd, signal, seed = 5)
hossa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_rec, rmse_ts_nd, signal, seed = 5)
save(ssa_err, hossa_err, file = paste0("./comp_results/cases_errors/two_periodics", full_case, ".RData"))
```

```{r}
print_errors <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> arrange(zapped_error, L_ssa, .by_group = TRUE)
}

summarise_errors <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> summarise(
    best = min(zapped_error),
    avg = mean(zapped_error),
    worst = max(zapped_error)
  )
}
```

```{r}
print_errors(ssa_err)
summarise_errors(ssa_err)
```

```{r}
print_errors(hossa_err)
summarise_errors(hossa_err)
```
